---
title: "USDA Data Science Training Program  \nIntermediate Assignment 3: Exploratory Data Analysis"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**Data:** USDA-ERS data on food environment factors (a subset of the Food Environment Atlas), along with county-level food security data from Feeding America. This assignment uses the variables listed below.

-   `State` - Name of the state
-   `County` - Name of the county
-   `CHILDPOVRATE15` - Child poverty rate, 2015
-   `PCT_LACCESS_HHNV15` - Percent of households with no car & low access to store, 2015
-   `METRO13` - Metro/nonmetro classification, 2013

For reference, a full list of the variables in this dataset is included at the end of the assignment.

**DataCamp:** The following DataCamp courses correspond to this exercise:

-   R: Exploratory Data Analysis in R, Introduction to Statistics in R, Reporting with R Markdown (bonus question)
-   Python: Exploratory Data Analysis in Python, Introduction to Statistics in Python, Building Dashboards with Dash and Plotly (bonus question)

**Assignment:**

*Part A: Load, inspect, and prep the data*

1.  Save `EDA_assignment_dataset.csv` to your computer and import the data into a dataframe called `food_env`.
2.  Print the summary statistics for all columns in `food_env`. Do any variables have missing data?
3.  Create a new dataframe named `food_env_full` that excludes rows with any missing data. How many rows were dropped?
4.  Convert the `METRO13` column to data type "logical."

```{r, echo=FALSE}
#A.1
food_env <- read.csv("EDA_assignment_dataset.csv")

#A.2
summary(food_env)

#A.3
food_env_full <- drop_na(food_env)
droppedRows <- as.numeric(nrow(food_env) - nrow(food_env_full))
print(paste("The number of dropped rows is", droppedRows))

#A.4
food_env_full$METRO13 <- as.logical(food_env_full$METRO13)
print(paste("The data type of METRO13 is", typeof(food_env_full$METRO13)))
``` 

*Part B: Explore the data*

5.  What are the minimum, maximum, mean, median, and standard deviation of `CHILDPOVRATE15`? Plot a histogram of child poverty rates to view the distribution.
6.  Create a density plot to examine the distribution of `PCT_LACCESS_HHNV15`. Does it look like the data are skewed? To investigate further, create a boxplot of the `PCT_LACCESS_HHNV15` data.
7.  Filter `food_env_full` to just counties in Virginia, Illinois, Michigan, Arkansas, and Ohio and save the smaller dataset as `food_env_5_states`.
8.  What are the minimum, maximum, mean, median, and standard deviation values of `CHILDPOVRATE15` and `PCT_LACCESS_HHNV15` in each state?
9.  Create faceted histograms or density plots of `CHILDPOVRATE15` and `PCT_LACCESS_HHNV15` by state. What do you notice about the distributions?
```{r, echo=FALSE}
#B.5
summary_stats <- function(data) {
  print(paste("Minimum:", min(data)))
  print(paste("Maximum:", max(data)))
  print(paste("Mean:", mean(data)))
  print(paste("Median:", median(data)))
  print(paste("Standard Deviation:", sd(data)))
}
print("Child poverty rate statistics")
print(summary_stats(food_env_full$CHILDPOVRATE15))

hist(food_env_full$CHILDPOVRATE15)

#B.6
ggplot(food_env_full, aes(x=PCT_LACCESS_HHNV15)) + geom_density()
boxplot(food_env_full$PCT_LACCESS_HHNV15)

#B.7
food_env_5_states <- food_env_full %>% filter(State %in% c("VA",
                                                        "IL",
                                                        "MI",
                                                        "AR",
                                                        "OH"))

#B.8

```


*Part C: Preliminary analysis*

10. Using `food_env_full`, compute the interquartile range of the `PCT_LACCESS_HHNV15` column to identify outliers. Remove rows with outliers from the data then create a new boxplot with the filtered dataframe. How does this boxplot compare to the one you created in question 6?
11. Make a scatterplot with `PCT_LACCESS_HHNV15` (outliers removed) on the x-axis and `CHILDPOVRATE15` on the y-axis. Add a straight line that shows the linear relationship between the two variables to your plot.
12. Compute the correlation between `PCT_LACCESS_HHNV15` and `CHILDPOVRATE15`. Do you think the relationship is strong? Are there other factors that could affect both the child poverty rate and a household's access to a grocery store in a county?

**Bonus:** Create a report in R Markdown or Dash using at least one of the plots you made, a table showing summary statistics for either `CHILDPOVRATE15` or `PCT_LACCESS_HHNV15`, and text explaining your analysis.

**Deliverables:**

-   Your code (the .R, .Rmd, .py, or .ipynb file).
-   If you've chosen to write your responses or to complete the bonus question in an R Markdown file, the knitted document with your responses.
-   These deliverables will be submitted through GitHub by the end of the program.

**Data Dictionary:**

-   `FIPS` - Numeric code that identifies the geographic area
-   `State` - Name of the state
-   `County` - Name of the county
-   `Pop2020` - Population of the county, 2020
-   `POVRATE15` - Poverty rate, 2015
-   `SNAPS17` - Count of SNAP-authorized stores in the county, 2017
-   `CHILDPOVRATE15` - Child poverty rate, 2015
-   `PCT_LACCESS_HHNV15` - Percent of households with no car & low access to store, 2015
-   `GROC16` - Count of grocery stores in the county, 2016
-   `FMRKT_SNAP18` - Count of Farmers' markets that report accepting SNAP, 2018
-   `METRO13` - Metro/nonmetro classification, 2013
-   `food_insecurity_2016` - Food insecurity rate, 2016
